Using cosmic variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-05-28 paper/ms.tex
removed comments from text

%In this \documentname we constrain the typical mass of dark matter halos
%hosting LAES at $z=3.1$. Our model is based on the number density
%information obtained in the recent large scale survey presented by XXX
%where XXX LAEs are detected over 7 fields of $\sim 46 \times
%35$Mpc$^{2}h^{-2}$ in area comoving in area corresponding to observed
%fields of $XXX$ deg$^{2}$.

%Both the spatial distribution and luminosities of the galaxies have,
%at least in a statistical sense, informatio to constrain theoretical
%model of LAEs. The most detailed theoretical models are also faced to
%diverse physical and astrophysical uncertainties in obtaining
%statistical prediction for the Lyman-alpha line. This uncertainties
%are largest impediment to consruct an ab-iitio model for LAEs.

%In this \documentname we want to step back and reduce the complexity of our
%model, with the sole objective of reproducing the cosmic variance in
%the number density of LAEs. Afterwards we will interpret the
%implications of this result for physical models for Lyman-alpha
%emitting galaxies.

%Our model is based on the predictions of a large volume high
%resolution N-body simulation describing the gravitational dynamics of
%dark matter. We do not have an strong bias towards the theoretical
%expectation of what the mass of the dark matter halo hosting the
%galaxy should be.  Instead, we fully explore the parameter space of
%our simplified model. The only cut we impose is that observed LAEs do
%not reside in dark matter halos with masses less than 10$^{10}$\hMsun
%[citation].

%In the following subsections we describe the most relevant features of
%the observational data, the N-body simulation we use, our model and
%its parameters together with the method to compare its predictions
%against observations.

@@

The observational benchamark we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey 
%Our observational reference are the recently published results of a
%panoramic survey of LAES at $z=3.1$ by \cite{Yamada2012}. 
%This survey
conducted with the Subaru 8.2m telescope and the Subaru Prime
Focus Camera, which has a field of view covering $34\times 27$ arcmin,
corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at
$z=3.09$. The narrow band filter is centered at $4977$ \AA with a
$77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and
$41$ Mpc $h^{-1}$ comoving scale for the detection of the
Lyman-$\alpha$ line centered at $z=3.09$. The authors report a total of
$2161$ LAEs with an observed equivalent width larger than $190$\AA
over a total survey area of $2.42$ deg$^{2}$, this corresponds to an
averaga surface number density of $0.20\pm 0.01$ arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs, this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs. The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs,
and the fild arotund the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs. 
%In Table 1 we summarize
%the values we use in throughout this paper for the each field, covered
%area, measured surface LAE number density and inferred number volume
%density.

@@

%For instance, having
%different selection criteria on the equivalent width (EW) can change
%the number of LAEs included in the catalog, yielding number densities
%different by a factor of two   
%different criteria on the equivalent width (EW) cuts to construct the
%LAE samples. Different cuts in the EW can change the number of LAEs to
%be included in the catalog. This cuts have an impact on the fainter
%LAEs which are more abundante than brighter ones. Different
%definitions of the EW cuts can yield number densities different by a
%factor of two [REF, I think Yamada has some numbers].

%Other
%surveys by XXX an XXX that cover similar regions, but they use
%different criteria on the equivalent width (EW) cuts to construct the
%LAE samples. Different cuts in the EW can change the number of LAEs to
%be included in the catalog. This cuts have an impact on the fainter
%LAEs which are more abundante than brighter ones. Different
%definitions of the EW cuts can yield number densities different by a
%factor of two [REF, I think Yamada has some numbers].

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking lenght of 0.17 times the interparticle
distance. 
%We have veryfied that the main results we present in this
%paper also hold if instead we use halo catalogs constructed from a the
%Bound Density Maxima (BDM) algorithm \citep{KlypinBDM} that are
%defined to have an density of 200 times the critical density. 
The minimum halo mass in the models we construct in this \documentname
correspond to groups of $\sim 75$ particles. The catalogs were
obtained from the publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
We focus our work on halos more massive than $1\times 10^{10}$\hMsun
that are resolved with at least $70$ particles.

@@

Next we construct mock surveys by making groups of $11$ mock fields
out of the $210$ available volumes. In total $15$ mock surveys are
constructed for each model $\mathcal{M}$. 
%mock fields into
%sets of $11$ different ways to
%construct the LAEs. The first way (match method) we
The grouping is done in two different ways. In the first way, called
{\texttt match}, we follow
the clustering of the observed fields. From the $11$ mock fields, $7$
are constructed from contigous fields in the simulation to mimic the
SSA22 region, $3$ are also contiguous between them but not to the first
$7$ fields to mimic the SXDS fields and finally $2$ non-contigous
fields to imitate the SDF and GOODS-North field. Our main goal with
this selection is to test the impact on the final statistics of having
$7$ clustered fields. The second way to group the mock fields is called {\texttt
  random}, whereby all the $11$ fields are selected in such a way as
to avoid that any two volumes are contigous.
%follow the observational setup and constructs $15$ different mock
%surveys, each one composed of $12$ mock fields, out of which $7$
%correspond to contiguous sub-boxes in the simulation to mimick the
%whole SSA22, $3$ are also contiguous between them but not to the first
%$7$ fields to mimick the SXDS fields and finally $2$ non-contiguous
%fields that correspond to the SDF and GOODS-North fields. This will
%produce $15$ different distributions for the number density for a
%given model ${\mathrm M}$. The second (random method) is similar to the first
%one. There aare $15$ different mock surveys with $12$ mock fields
%each, but this time each field corresponds to uncorrelated sub-boxes
%in the simulation. The third (full method) way in only has $1$ mock survey
%containing all the $210$ mock fields, in this setup there is only one
%predicted number density distribution for each model ${\mathrm M}$ 
%We create the mock surveys in two different ways. First we follow 
%The advantage of these three sampling ways is that they allow us to
%explore the effects of both cosmic variance and the correlation
%between fields. Comparing the results of the first and second method
%will help us to quantify the effect of field correlation, while
%compating the first and the third method will serve us to gauge the
%impace of cosmic variance.

@@

%For each mock survey generated in a given model ${\mathcal M}$ we compute the LAE surface density
%distributions for the three different ways of grouping the mock
%fields, as described in the previous section. For each sub-volume we
%project the positions of the LAE hosting halos along the $z$ direction
%and calculate its surface number density in units of sources per
%arcmin$^{2}$. For each number density distribution we perform a
%Kolmogorov-Smirnov against the $12$ surface density observational
%values. From this test we obtain the value $0<P<1$ to reject the null
%hypothesis, namely that the two data sets come from the same
%distribution. In this paper we use values of $P>0.1$ to consider that
%the simulated and observed number densities come from the same
%distribution.

%We generate a series of models ${\mathcal M}$ with different input
%parameters$\{M_{\rm min}, M_{\rm max}, f_{\rm occ}\}$ as
%follows. $M_{\rm min}$ and $M_{\rm max}$ are allowed to take 30
%different values evenly spaced by $0.1$ dex, $M_{\rm min}$ ranges from
%$\log_{10}M_{\rm min}=10.0$ up to $\log_{10}M_{\rm min}=12.9$, while
%$M_{\rm max}$ range  from $\log_{10}M_{\rm min}=10.1$ up to
%$\log_{10}M_{\rm min}=13.0$. The occupation fraction $f_{\rm occ}$
%takes 100 different values from $0.01$ to $1.00$ regularly spaced by
%$0.01$. In total the number of different sets of input parameters to
%be explored is $30 \times 30\times 100 = 9\times 10^{4}$.

@@

%This basic result provides the basis to understand the range of models
%${\mathcal M}$ expected trends for the
%LAEs' preferred mass and the occupation fraction.  
%From this Figure we
%can read which models do not have achance reproduce the
%observations. Regions in the plot where the halo surface density
%values are below the observational constraint correspond to high
%masses halo masses. For a LAE model ${\mathcal M}$ with a minimum mass
%$M_{\rm   min}> 3 \times 10^{11}$\hMsun located inthat mass range, the
%surface density is too low compared with observations.

@@

%The raw results of our experiments yield a prefered range for
%halo masses hosting LAES bounded by a minimum mass $10.5<\log_{10}M_{\rm
%  min}<11.5$ and without any limitation on the  maximum
%mass. In this regime the average occupation fraction tends to increase
%with $M_{\rm min}$. These results discard models that were
%nevertheless disfavored from the very beginning based on the results
%of the mass functions (see Figure 1.).

%Figure \label{figure:landscape} presents the regions in the parameter
%space $M_{\rm min}$-$M_{\rm max}$ where the KS test yields values of
%$P>0.05$. We consider that for those models it is not possible to rule out the null
%hypothesis, namely that the number density in simulated data and the observations come
%from the same parent distribution. Each panel corresponds to the three different ways of
%grouping the mock fields. In the case of the methods {\tt{Match}}
%and {\tt{Random}} the color code indicates the fraction of these 15
%mock surveys with with $P>0.1$. The third panel shows the result for
%the method {\tt{Full}}, in this case the color code correspond
%to the value maximum value of $100\times P$ for a model with those
%mass ranges.

%These results clearly distinguish three mass regimes. In the first regime, at
%high mass values, we find that LAE models with minimum mass of $M_{\rm
%  min}>10^{11.5}$\hMsun are not compatible with observations. There is
%a second regime for masses below $M_{\rm
%  min}=3\times 10^{10}$ any values for $M_{\rm min}$ and $M_{\rm max}$
%can be made compatible with observations, provided that
%$f_{\rm occ}$ is fine tuned to do it. In an intermediate mass regime,
%for minimum mass values $3\times 10^{10}\hMsun < M_{\rm min}< 3\times 10^{11}\hMsun$ only a
%limited range of models with $M_{\rm max}$ with occupation fraction
%$f_{\rm occ}\sim 1$ is able to reproduce observations.

%In these three different mass regimes the occupation monotonically
%decreases as a function of the minimum halo mass $M_{\rm min}$. In
%Figure XXX we show this treend in three panels following the same
%correspondence as Figure XXX. From these results we interpret that the
%different mass regimes that were identified correspond to best fit
%models with $f_{\rm occ}\sim 1$, $0<f_{\rm occ}<1$ and $f_{\rm
%  occ}\sim 0$, respectively.

%The two {\tt match} and {\tt random} methods present the highest
%number of matching mock surveys in the medium mass regime. However, it
%is important to keep in mind that not all the mock surveys for a
%sucessful model ${\mathrm M}$ present a high value $P>0.1$, only a
%modest fraction seems to be consistent with observations. This shows that the
%cosmic variance is still present on the physical scales probed by
%observations. This will be considered in more detail in the discussion
%seccion.

%To illustrate this point, in Figure \ref{fig:mocks} we present the
%results for two mock fields for the {\tt{Match}} method for a model
%with the same parameters, but two extreme values for the KS test.

%Conversely there are different models where the KS test yield values of
%$P\sim 1$. To illustrate the kind of success represented by these
%models, we have selected these best ones in the case of the method
%{\tt{MatchObs}}. Figure XXX shows in the main panel the spatial
%distribution of the mock surveys, the smaller panel shows the
%corresponding surface density distribution and the observational
%constraint.

%In what follows we will focus our discussion on the mocks constructed
%with the {\tt{Match}} and {\tt{Random}} methods.

@@

%In this section we consider three different ways to impose
%tighter constraints on this mass range by making use of the results we
%have derived so far together with additional statistical and
%astrophysical constraints. In the first constraint we select the
%models where all the mock surveys present KS-test values consistent
%with observations. The second constraints uses recent observational
%results on the average occupation fraction for LAEs at high-z. The
%third exploits the information in the Angular Correlation Function
%(ACF).

@@

Figure \ref{fig:high_success_rate} shows the number of models
that have at least $N_{\rm high-P}$ mocks with $P>0.05$ for the {\texttt
  match} and {\texttt random} methods.  This shows that there are
here are between $80$ to $100$ models with all the 15 realizations with
$P>0.05$. This a reduction of a factor of $\sim 6$ with repect to the
total number of mocks with at least one consistent mock. 
%Here we focuse on the latter models. 
%The best models
%represent $\sim 15\%$ of the number of initially considered good
%models.

@@

We limit the occupation fraction to
be in the range 
%further pick models that have an occupation fract
$f_{\rm
  occ}\leq 0.2$ to allow for some flexibility on the time evolution
and the uncertainty in the dust correction used to infer the escape fraction. 
%Considering an occupation fraction $f_{\rm occ}=0.20$
%another $39$This constraint help us to
%select $\sim 10\%$ of theoriginal models that were considered as
%consistent with observations in the previous section. 
Figure  \ref{fig:restriction_mock_and_f_occ}  shows the prefered
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm
  min}-f_{\rm occ}$ for the {\tt match} and {\tt   random} methods.
With this additional constraint between $35$ to $40$ models are
consistent with observations, this represents a factor of $\sim 2$
reduction. All the models with $\log_{10}M_{\rm min}\geq 11.0$ are now
excluded and the best models are now clustered around a narrow region
in parameter space. The list for the model parameters is in the
Appendix in Table \ref{table:models_match}.

%In terms of the constraints done in the previous sub-section, we find
%that the models in this region of parameter space have an average
%number of XX mock catalogs consistent with observations.

@@

lun abr 29 07:38:58 COT 2013

>>>>

@@

The observational benchamark we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey  conducted with
the Subaru 8.2m telescope and the Subaru Prime Focus Camera, which has
a field of view covering $34\times 27$ arcmin, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$. The narrow
band filter is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors report a total  $2161$ LAEs with an
observed equivalent width larger than $190$\ over a total survey area
of $2.42$ deg$^{2}$, this corresponds to  averaga surface number
density of $0.20\pm 0.01$ arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs, this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs. The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs,
and the fild arotund the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs.

@@

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking lenght of 0.17 times the interparticle
distance. The minimum halo mass in the models we construct in this \documentname
correspond to groups of $\sim 75$ particles. The catalogs were
obtained from the publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
We focus our work on halos more massive than $1\times 10^{10}$\hMsun
that are resolved with at least $70$ particles.

@@

Next we construct mock surveys by making groups of $11$ mock fields
out of the $210$ available volumes. In total $15$ mock surveys are
constructed for each model $\mathcal{M}$. The grouping is done in two
different ways. In the first way, called {\texttt match}, we follow
the clustering of the observed fields. From the $11$ mock fields, $7$
are constructed from contigous fields in the simulation to mimic the
SSA22 region, $3$ are also contiguous between them but not to the first
$7$ fields to mimic the SXDS fields and finally $2$ non-contigous
fields to imitate the SDF and GOODS-North field. Our main goal with
this selection is to test the impact on the final statistics of having
$7$ clustered fields. The second way to group the mock fields is called {\texttt
  random}, whereby all the $11$ fields are selected in such a way as
to avoid that any two volumes are contigous.

@@

@@

@@

@@

@@

Figure \ref{fig:high_success_rate} shows the number of models
that have at least $N_{\rm high-P}$ mocks with $P>0.05$ for the {\texttt
  match} and {\texttt random} methods.  This shows that there are
here are between $80$ to $100$ models with all the 15 realizations with
$P>0.05$. This a reduction of a factor of $\sim 6$ with repect to the
total number of mocks with at least one consistent mock.

@@

We limit the occupation fraction to
be in the range $f_{\rm  occ}\leq 0.2$ to allow for some flexibility on the time evolution
and the uncertainty in the dust correction used to infer the escape fraction. 
Figure  \ref{fig:restriction_mock_and_f_occ}  shows the prefered
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm
  min}-f_{\rm occ}$ for the {\tt match} and {\tt   random} methods.
With this additional constraint between $35$ to $40$ models are
consistent with observations, this represents a factor of $\sim 2$
reduction. All the models with $\log_{10}M_{\rm min}\geq 11.0$ are now
excluded and the best models are now clustered around a narrow region
in parameter space. The list for the model parameters is in the
Appendix in Table \ref{table:models_match}.

@@
